---
title: "asn2139_hw2"
author: "Akanksha Nalatwad"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)

```
## Problem 1

Importing Dataset

```{r}
trash_df=
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet="Mr. Trash Wheel",
  range=cell_cols("A:N")) %>% 
      janitor::clean_names () %>% 
     drop_na(dumpster) %>% 
    mutate(
      sports_balls= round(sports_balls),
      sports_balls= as.integer(sports_balls)
    )
```

Read preciptation data for 2017 and 2018

```{r}
precip_2018=
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet="2018 Precipitation",
             skip=1) %>% 
            janitor::clean_names () %>% 
            drop_na(month) %>% 
            mutate(year=2018) %>% 
            relocate(year)

precip_2017=
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet="2017 Precipitation",
             skip=1
             ) %>% 
            janitor::clean_names() %>% 
            drop_na(month) %>% 
            mutate(year=2017) %>% 
            relocate(year)
  
          

```

```{r}
month_df=
      tibble(
            month=1:12,
            month_name=month.name
      )
        
precip_df=bind_rows(precip_2018, precip_2017)

precip_dif= left_join(precip_df, month_df, by="month")
      
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trash_df)` rows in our final dataset. Additional data sheets include month precipitation data. In this dataset:

* The median number of sports balls found in a dumpster in 2017 was `r trash_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.

## Problem 2



```{r}
  
transit_df=
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select (line, station_name, station_latitude, station_longitude, starts_with("route"), entry, vending, entrance_type, ada) %>% 
  mutate(entry=ifelse(entry=="YES",TRUE,FALSE)) %>% 
  mutate(vending=ifelse(vending=="YES", TRUE, FALSE))

```

## Description of Dataset

There are `r ncol(transit_df)` characteristics describing the `r nrow(transit_df)` entrys and exits of subway stations in NYC.
The lines includeed in this dataset are `r unique (pull(transit_df, line))`.The routes in this data set include `r unique (pull(transit_df,route1))`. The types of entrances described are `r unique(pull(transit_df,entrance_type))`.


Answer the following questions using these data:

How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St A/B/C/D; 125st 1; 125st 4/5); the distinct function may be useful here.

```{r}
distinct(transit_df,station_name)
```
There are 356 station names in this dataset. 

How many stations are ADA compliant?

```{r}

var(transit_df)
ada_compliant=
select(transit_df, ada)

  summary(ada_compliant)
```
468 stations are ADA compliant.

What proportion of station entrances / exits without vending allow entrance?

```{r}
no_vending_df=
filter(transit_df,vending=="NO")


no_vending_entry_df=
filter(no_vending_df,entry=="YES" )

(nrow(no_vending_entry_df))/(nrow(no_vending_df))

```

The proportion is `r(nrow(no_vending_entry_df))/(nrow(no_vending_df))`. 


Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?



