---
title: "asn2139_hw2"
author: "Akanksha Nalatwad"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)

```
## Problem 1

Importing Dataset

```{r}
trash_df=
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet="Mr. Trash Wheel",
  range=cell_cols("A:N")) %>% 
      janitor::clean_names () %>% 
     drop_na(dumpster) %>% 
    mutate(
      sports_balls= round(sports_balls),
      sports_balls= as.integer(sports_balls)
    )
```

Read preciptation data for 2017 and 2018

```{r}
precip_2018=
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet="2018 Precipitation",
             skip=1) %>% 
            janitor::clean_names () %>% 
            drop_na(month) %>% 
            mutate(year=2018) %>% 
            relocate(year)

precip_2017=
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet="2017 Precipitation",
             skip=1
             ) %>% 
            janitor::clean_names() %>% 
            drop_na(month) %>% 
            mutate(year=2017) %>% 
            relocate(year)
  
          

```

```{r}
month_df=
      tibble(
            month=1:12,
            month_name=month.name
      )
        
precip_df=bind_rows(precip_2018, precip_2017)

precip_dif= left_join(precip_df, month_df, by="month")
      
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trash_df)` rows in our final dataset. Additional data sheets include month precipitation data. In this dataset:

* The median number of sports balls found in a dumpster in 2017 was `r trash_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.

## Problem 2



```{r}
  
transit_df=
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select (line, station_name, station_latitude, station_longitude, starts_with("route"), entry, vending, entrance_type, ada) %>% 
  mutate(entry=ifelse(entry=="YES",TRUE,FALSE)) %>% 
  mutate(vending=ifelse(vending=="YES", TRUE, FALSE))

```

## Description of Dataset

There are `r ncol(transit_df)` characteristics describing the `r nrow(transit_df)` entrys and exits of subway stations in NYC.
The lines includeed in this dataset are `r unique (pull(transit_df, line))`.The routes in this data set include `r unique (pull(transit_df,route1))`. The types of entrances described are `r unique(pull(transit_df,entrance_type))`. The data cleaning has involved janitor clean names, narrowing down the varaibles that we are interested in, and making the variables "entry" and "vending" clear. This data is not tidy.


Answer the following questions using these data:

How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St A/B/C/D; 125st 1; 125st 4/5); the distinct function may be useful here.

```{r}
nrow(distinct(transit_df,station_name,line))
```
There are `r nrow(distinct(transit_df,station_name,line))` station names in this dataset. 

How many stations are ADA compliant?

```{r}

var(transit_df)
ada_compliant=
select(transit_df, ada)

  summary(ada_compliant)
  
```
468 stations are ADA compliant.

What proportion of station entrances / exits without vending allow entrance?

```{r}

no_vending_df=
select(transit_df,vending,entry) %>% 
filter(vending=="FALSE") %>%  
summary(no_vending_df)

```

The proportion is `r 69/183`. 


Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?

```{r}
tidy_transit_df=
mutate(transit_df,route8=as.character(route8),route9=as.character(route9),
                  route10=as.character(route10),route11=as.character(route11)) %>% 
pivot_longer(route1:route2:route3:route4:route5:route6:route7:route8:route9:route10:route11, 
             values_to="route_name") 
```


##Problem 3

```{r}

```

Loading, cleaning, and making changes to pols_df
```{r}
pols_df=
  read_csv("./data/pols-month.csv") %>% 
   janitor::clean_names() %>% 
   separate("mon", c("year", "month", "day"), sep="-") %>% 
    mutate(month =recode(month,'01'="jan",'02'="feb", '03'="mar", '04'="apr",'05'="may",
                               '06'="jun",'07'="jul",'08'="aug",'09'="sept", '10'="oct",
                               '11'="nov",'12'="dec")) %>% 
    mutate(prez_gop = recode(prez_gop, "1"="gop","0"="dem")) %>% 
    rename(president=prez_gop) %>% 
    select(-prez_dem,-day) 
```

```{r}
snp_df=
  read_csv("./data/snp.csv") %>% 
  janitor::clean_names() %>% 
   separate("date", c("month", "day", "year"), sep="/") %>% 
    mutate(month =recode(month,'01'="jan",'02'="feb", '03'="mar", '04'="apr",'05'="may",
                               '06'="jun",'07'="jul",'08'="aug",'09'="sept", '10'="oct",
                               '11'="nov",'12'="dec")) %>% 
    relocate(year,month) %>% 
    select(-day) 
```

import and tidy umemployment dataset

```{r}

unemploy_df=
  read_csv("./data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(jan:feb:mar:apr:may:jun:jul:aug:sep:oct:nov:dec,
               values_to= "unemployed") %>% 
  rename(month=name) %>% 
  mutate(year=as.character(month))

```
joining the datasets

```{r}
compiled_df=
left_join(pols_df,snp_df, by=c("month","year")) 

left_join(compiled_df,snp_df, by=c("month","year")) 
```

